\stktokens{}: Enforcing Well-Bracketed Control Flow and Stack Encapsulation Using Linear Capabilities

### 2019-07-22 popl-2019-journal-v/paper.tex
[POPLjv] Added more explanation to code memory padding

%TODO distinguish closure seals from return seals.
The $\wdjud[\ta]{\comp}$ judgement has two rules: \textsc{Main} and \textsc{Base}.
The \textsc{Main} rule requires the component to have an entry point in terms of a main pair from the exports.
Further, the base component should satisfy the well-formedness judgement.
The \textsc{Base} judgement ensures sure that the component has a bit of structure.
Roughly speaking, the \textsc{Base} rule requires the following:
\begin{itemize}
\item \textit{Code and data memory are disjoint.}
\item \textit{Code memory is padded with zeros ($\mspad$).} This prevents the memory of two code memories from being spliced together.\todo[inline]{Dominique: more explanation needed?}
\item \textit{Data address space can be partitioned in a linear part $A_\var{linear}$ and a non-linear part $A_\var{non-linear}$.}
  The component judgement ensure that the linear address space of a component is at most referenced by one capability. This means that any capability for the linear address space is unique.
\item \textit{There is a partition $A_\var{own}$ of $A_\var{linear}$ that assigns a set of linear-addressed to each address of data memory.} The partitions are used by the component word judgement ($\vdash_\mathrm{comp-word}$) which ensures linear capabilities only address the addresses in the linear address space that is owned by the address where the capability resides in memory.
  \todo[inline]{Maybe it would be clearer to merge this point and the previous point and explain that all memory can either be addressed by any number of non-linear caps or by at most one linear capability in the component's data memory.}
\item \textit{All import addresses are part of the data memory.} 
\item \textit{Import and export symbols are disjoint.}
\item \textit{One of the following two are true}
  \begin{itemize}
  \item \textit{The code address space is disjoint from the trusted address space $\ta$ and there are no return seals.} In this case, the component contains untrusted code. We are interested in WBCF and LSE from the perspective of the trusted code, so we do not let untrusted memory have return seals\footnote{Untrusted code still has access to seals that could be used to protect calls.}.
  \item \textit{The code address space is part of the trusted address space $\ta$.} In this case, the component contains trusted code. We do not impose requirements on the return seals here (the auxiliary judgements will impose restrictions on them), so the trusted component can have the return seals necessary for its calls.
  \end{itemize}
\item \emph{The data address space is disjoint from the trusted address space $\ta$.} Data memory is not executable, so the trusted addresses never include the data memory addresses.
\item \textit{The code memory satisfies the component-code judgement (\figurename~\ref{fig:well-formed-code}).}
\item \textit{Each word in the data memory satisfies the component-word judgement (\figurename~\ref{fig:well-formed-word}).} This must be with respect to the linear addresses assigned to this address.
\item \textit{All the exports satisfy the components-export judgement (\figurename~\ref{fig:well-formed-export}).}
\end{itemize}

>>>>

%TODO distinguish closure seals from return seals.
The $\wdjud[\ta]{\comp}$ judgement has two rules: \textsc{Main} and \textsc{Base}.
The \textsc{Main} rule requires the component to have an entry point in terms of a main pair from the exports.
Further, the base component should satisfy the well-formedness judgement.
The \textsc{Base} judgement ensures sure that the component has a bit of structure.
Roughly speaking, the \textsc{Base} rule requires the following:
\begin{itemize}
\item \textit{Code and data memory are disjoint.}
\item \textit{Code memory is padded with zeros ($\mspad$).}
  This prevents code memories from being spliced together.
  If the capabilities for two code memories can be spliced together, then the execution of one code memory can continue into the other creating an unintended control-flow.
  Further, the possible control-flows of a component suddenly depends on the memory locations of all the components\footnote{Alternatively to this syntactic condition on components, we could require trusted components to never splice executable capabilities of unknown origin.}.
\item \textit{Data address space can be partitioned in a linear part $A_\var{linear}$ and a non-linear part $A_\var{non-linear}$.}
  The component judgement ensure that the linear address space of a component is at most referenced by one capability. This means that any capability for the linear address space is unique.
\item \textit{There is a partition $A_\var{own}$ of $A_\var{linear}$ that assigns a set of linear-addressed to each address of data memory.} The partitions are used by the component word judgement ($\vdash_\mathrm{comp-word}$) which ensures linear capabilities only address the addresses in the linear address space that is owned by the address where the capability resides in memory.
  \todo[inline]{Maybe it would be clearer to merge this point and the previous point and explain that all memory can either be addressed by any number of non-linear caps or by at most one linear capability in the component's data memory.}
\item \textit{All import addresses are part of the data memory.} 
\item \textit{Import and export symbols are disjoint.}
\item \textit{One of the following two are true}
  \begin{itemize}
  \item \textit{The code address space is disjoint from the trusted address space $\ta$ and there are no return seals.} In this case, the component contains untrusted code. We are interested in WBCF and LSE from the perspective of the trusted code, so we do not let untrusted memory have return seals\footnote{Untrusted code still has access to seals that could be used to protect calls.}.
  \item \textit{The code address space is part of the trusted address space $\ta$.} In this case, the component contains trusted code. We do not impose requirements on the return seals here (the auxiliary judgements will impose restrictions on them), so the trusted component can have the return seals necessary for its calls.
  \end{itemize}
\item \emph{The data address space is disjoint from the trusted address space $\ta$.} Data memory is not executable, so the trusted addresses never include the data memory addresses.
\item \textit{The code memory satisfies the component-code judgement (\figurename~\ref{fig:well-formed-code}).}
\item \textit{Each word in the data memory satisfies the component-word judgement (\figurename~\ref{fig:well-formed-word}).} This must be with respect to the linear addresses assigned to this address.
\item \textit{All the exports satisfy the components-export judgement (\figurename~\ref{fig:well-formed-export}).}
\end{itemize}
